---
title: "Mortality in SA"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
# Setup code runs once when start the app

# Setup libraries and data
library(flexdashboard)
library(shiny)
library(ggplot2)
library(readr)
library(dplyr)
library(tidyr)
library(leaflet)
# Note: this data needs to be in local repo since needs to be read in.
mortality_zaf <- read_csv("~/Data science/R/shiny-tutorial/mortality_zaf.csv")
population_zaf <- read_csv("~/Data science/R/shiny-tutorial/population_zaf.csv")

# create required dataframe for charts
#totals
totals <- population_zaf %>%
  filter(year == 2013) %>%
  select(province, total)
#spread of data
# this is the gplyr and tidyr pipe
compare_provinces <- mortality_zaf %>%
  left_join(totals) %>%
  mutate(rate = deaths / total * 1e3) %>%
  group_by(province, indicator) %>%
  summarise(rate = mean(rate)) %>%
  ungroup %>%
  left_join(totals) %>%
  spread(indicator, rate) # not including the data argument because included through the pipe %>%

# remove all causes indicator from mortality dataset, after created compare_provinces
mortality_zaf <- mortality_zaf %>%
  filter(indicator != "All causes")



```

Column {.sidebar}
=======================================================================
__About the sidebar:__ For the global sidebar, use level 1 header "=" underneath.
Now this sidebar will be visible all the time. 'Column"" is a key word for the sidebar setup.

__About the dashboard:__ Data from Code4SA mortality report.


__About the charts:__ Use the input below to select what you want to explore.

```{r sidebarinput}
# shiny input in sidebar for user options
selectInput(
  inputId = "indicator",
  label = "Cause of death:",
  choices = unique(mortality_zaf$indicator),
  selected = "Non-natural causes"
)


```

Map {data-icon="fa-bullseye"} 
==================================================

### Where is the mortality rate highest?

```{r mapPage}
# create reactive expression for data subset that will be used across all pages
selected_df <- reactive({
  subset_df <- compare_provinces[, c(1,
                        which(colnames(compare_provinces) == input$indicator)
                        )]
  colnames(subset_df) <- c("province", "indicator") 
  population_zaf %>%
    filter(year == 2013) %>%
    left_join(subset_df) %>%
    mutate(indicator = indicator / sum(indicator, na.rm = TRUE))
  #subset_df
})

renderLeaflet({
  leaflet(selected_df()) %>%
    addProviderTiles("CartoDB.Positron") %>% #or leave blank for default tiles
    addCircles(lng = ~longitude,
               lat = ~latitude,
               radius = ~sqrt(indicator)*1e5,
               weight = 2.5,
               color = "red",
               popup = ~province)
  
})
```


Comparing provinces {data-icon="fa-birthday-cake"}
===============================================

### How does the selected mortality rate compare to the overall mortality rate?

```{r provinceComparison}
scatter_df <- reactive({
  subset_df <- compare_provinces[, c(1,2,
                        which(colnames(compare_provinces) == "All causes"),
                              which(colnames(compare_provinces) == input$indicator))]
  colnames(subset_df) <- c("province", "total", "x", "y")
  subset_df
  
})

renderPlot({
  ggplot(scatter_df(), 
       aes(x, y, 
           label = province
           )) + 
  geom_point(aes(size = total),
             alpha = 0.8, color = "red") +
  geom_text(vjust = -2.5) +
    theme_minimal() +
    labs(x = "All causes",
         y = input$indicator)
})
```


Changes in time {data-icon="fa-anchor"}
==================================================

### How have the number of deaths changed over time?

```{r linechart}
renderPlot({
  mortality_zaf %>%
    filter(indicator == input$indicator) %>%
    ggplot()
})
```


