---
title: "Shiny with R Markdown"
author: "Bryony"
date: "February 17, 2017"
runtime: shiny
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)#, echo = FALSE, warningg = FALSE, message = FALSE)
```


```{r libraries} 
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(shiny)
# Note: this data needs to be in local repo since needs to be read in.
mortality_zaf <- read_csv("~/Data science/R/shiny-tutorial/mortality_zaf.csv")
population_zaf <- read_csv("~/Data science/R/shiny-tutorial/population_zaf.csv")
```


## Sorting out the data
```{r data}
totals <- population_zaf %>%
  filter(year == 2013) %>%
  select(province, total)

# this is the gplyr and tidyr pipe
compare_provinces <- mortality_zaf %>%
  left_join(totals) %>%
  mutate(rate = deaths / total * 1e3) %>%
  group_by(province, indicator) %>%
  summarise(rate = mean(rate)) %>%
  ungroup %>%
  left_join(totals) %>%
  spread(indicator, rate) # not including the data argument because included through the pipe %>%
  
```


## Making the plot - a regular static plot
```{r staticplot}
ggplot(compare_provinces, 
       aes(`Tuberculosis (A15-A19)`,
           `Human immunodeficiency virus [HIV] disease (B20-B24)`, label = province
           )) + 
  geom_point(aes(size = total),
             alpha = 0.8, color = "midnightblue") +
  geom_text(vjust = -1.5) +
    theme_minimal()

```


## Using shiny stuff in Rmd: creating inputs
When using shiny in Rmd, there is no code for the sidebar layout, frame etc, just the objects
```{r shinyinputs}
selectInput(inputId = "x_axis",
            label = "Choose the horizontal axis indicator:",
            choices = colnames(compare_provinces)[3:20],
            selected = "Non-natural causes"
            )

selectInput(inputId = "y_axis",
            label = "Choose the vertical axis indicator:",
            choices = colnames(compare_provinces)[3:20],
            selected = "Tuberculosis (A15-A19)"
            )

```


## Making a reactive expression
Note: the curly braces "{}" are visual cue for reactive code
```{r reactive}
selected_df <- reactive({
  subset_df <- compare_provinces[, c(1,2,
                        which(colnames(compare_provinces) == input$x_axis), 
                        which(colnames(compare_provinces) == input$y_axis))]
  colnames(subset_df) <- c("province", "total", "x", "y")
  subset_df
})
```


## Making the plot using a reactive expression
Note: this aspect of the code would sit within the server function.  In Rmd just put things in sequence.
```{r shinyplot}
renderPlot({
  ggplot(selected_df(), 
       aes(x, y, 
           label = province
           )) + 
  geom_point(aes(size = total),
             alpha = 0.8, color = "magenta") +
  geom_text(vjust = -1.5) +
    theme_minimal() +
    labs(x = input$x_axis,
         y = input$y_axis)
})
```


## Making the plot a specficied shape...
Note: this aspect of the code would sit within the server function.  In Rmd just put things in sequence.
```{r shinyplotincolumn}
fillCol(height = 800,
        renderPlot({
        ggplot(selected_df(),
        aes(x, y,
        label = province)) +
        geom_point(aes(size = total),
        alpha = 0.8, color = "magenta") +
        geom_text(vjust = -1.5) +
        theme_minimal() +
        labs(x = input$x_axis,
        y = input$y_axis)
        }, height = 800))
```




